Supplement to Automatic Matching of Bullet Lands

### 2016-04-05 imaging-paper.Rnw
Fix all the build issues

\address{Address of the Authors\\
Department of Statistics and Statistical Laboratory\\ Snedecor Hall\\
Ames Iowa 50011-1210\\
\printead{e1}\\
\phantom{E-mail:\ }\printead*{e2}\\
\phantom{E-mail:\ }\printead*{e3}}
\end{aug}
%
\begin{abstract}
In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel. During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet. Identifying these striation markings in an attempt to match two bullets is the primary goal. We propose an automated framework for the analysis of the 3D surface measurements, that first transcribes the markings into a 2D plotting framework. This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels. The automatic matching routine we propose manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
\end{abstract}

@@

options(replace.assign=TRUE,scipen=3, digits=2)
bstats <- read.csv("data-25-25/bullet-stats.csv")

@@

lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)

@@

\subsection{Impact of bullet height}\label{sec:heights}
The height at which signatures are extracted for a comparison between bullet lands matters -- signatures taken from heights that are further apart, show more pronounced differences between the signatures. 
This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of the height that was used in a matching. Visually, matches degrade if the signatures upon which the match is based are from heights further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion).
However, we can extract signatures from multiple heights of the same bullet land for an initial assessment of its quality. 
This situation is assessed and discussed in more detail below:

@@

#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
#  lines <- striation_identify(lofX, threshold = threshold)
  peaks1$lines$bullet <- b12[1]
  peaks2$lines$bullet <- b12[2]
  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
             lop[[4]], lop[[5]], lop[[6]],
             ncol=2)
@
\caption{\label{fig:crosscuts2}Signatures  for barrel 3, bullet~1-1 extracted from varying heights. Initially, the match between signatures taken at heights 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet. At a  level of $175\mu m$ the bullet's signature stabilizes. For this land, matches should not be attempted at lower heights. }
\end{figure}
%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
by comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures. In the approach here, we keep increasing the height $x$ at which the signature is taken until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
`stability' is  defined as two aligned signatures from heights chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.

@@

Figure~\ref{fig:varysmooth} displays the  peaks and valleys detected in the same signature at smoothing factors of 5, 25, and 45, respectively. The dark line corresponds to the smoothed values, while the grey line in the back shows the raw signature. The choice of smoothing factor is a classical decision of a bias/variance trade-off. It is immediately clear that a small smoothing factor like 5 is a poor choice. It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected. Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing. Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artifacts in the analysis.

@@

Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches. Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$). %The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.
All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches. The predictive power of each one of these features is shown in the form of the Receiving Operating Characteristic (ROC) curves in Figure~\ref{fig:rocs}. The features are arranged in descending order according to the area under the curve (AUC). The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys. The maximal number of CMS is only in the seventh position here. The overall high AUC values indicate, that we can successfully employ machine learning methods to distinguish matches from non-matches.
%Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3D topological surface measurements. There are quite large differences in these averages, indicating

Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files. The resulting tree is shown in Figure~\ref{fig:tree}. A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} lands is being matched correctly. Interestingly, the number of consecutive matching striae does not feature in this evaluation. 
Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree. Aside from  cross correlation, the total number of matches is also included in the decision rule. 
Between cross-correlation and CMS, cross-correlation has higher  predictive power. This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2D microscopic images of bullets.
Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure. The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}. However, in this study authors did not consider the number of matches and non-matches.

\begin{figure}[hbtp]
  \centering
<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\textwidth'>>=
library(rpart)
library(rpart.plot)
library(RColorBrewer)
vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
names(bstats)[9:10] <- c("#matches", "#non-matches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
cc <- read.csv("csvs/crosscuts-25-new.csv")
# cc$bullet <- gsub(".*//","",cc$path)
# cc$bullet <- gsub(".x3p", "", cc$bullet)
# bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
#                   bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]

@@

The feasibility of creating a database of ballistic images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks. This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system. A further re-assessment by \citet{deceuster:2015} came to the same conclusions based on the EvoFinder system. 
The NRC report also found, that results from 2D images can be improved when matches are based on 3D images. This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the total number of matches and mismatches are available for a match based on 2D features.

>>>>

\address{Address of the Authors\\
Department of Statistics and Statistical Laboratory\\ Snedecor Hall\\
Ames Iowa 50011-1210\\
\printead{e1}\\
\phantom{E-mail:\ }\printead*{e2}\\
\phantom{E-mail:\ }\printead*{e3}}
\end{aug}
%
\begin{abstract}
In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel. During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet. Identifying these striation markings in an attempt to match two bullets is the primary goal. We propose an automated framework for the analysis of the 3D surface measurements, that first transcribes the markings into a 2D plotting framework. This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels. The automatic matching routine we propose manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
\end{abstract}

@@

options(replace.assign=TRUE,scipen=3, digits=2)
bstats <- read.csv("data-25-25/bullet-stats-old.csv")

@@

lines <- striation_identify(peaks1$lines, peaks2$lines)

@@

\subsection{Impact of bullet height}\label{sec:heights}
The height at which signatures are extracted for a comparison between bullet lands matters -- signatures taken from heights that are further apart, show more pronounced differences between the signatures. 
This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of the height that was used in a matching. Visually, matches degrade if the signatures upon which the match is based are from heights further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion). However, we can extract signatures from multiple heights of the same bullet land for an initial assessment of its quality. By comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures. In the approach here, we keep increasing the height $x$ at which the signature is taken until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where `stability' is  defined as two aligned signatures from heights chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.

@@

#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
#  lines <- striation_identify(lofX, threshold = threshold)
  peaks1$lines$bullet <- b12[1]
  peaks2$lines$bullet <- b12[2]
  lines <- striation_identify(peaks1$lines, peaks2$lines)

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
             lop[[4]], lop[[5]], lop[[6]],
             ncol=2)
@
\caption{\label{fig:crosscuts2}Signatures  for barrel 3, bullet~1-1 extracted from varying heights. Initially, the match between signatures taken at heights 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet. At a  level of $175\mu m$ the bullet's signature stabilizes. For this land, matches should not be attempted at lower heights. }
\end{figure}
%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:

@@

Figure~\ref{fig:varysmooth} displays the  peaks and valleys detected in the same signature at smoothing factors of 5, 25, and 45, respectively. The dark line corresponds to the smoothed values, while the grey line in the back shows the raw signature. The choice of smoothing factor is a classical decision of a bias/variance trade-off. It is immediately clear that a small smoothing factor like 5 is a poor choice. It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected. Given that striation widths are typically much larger, we are in effect muddying the waters by performing such minimal smoothing. Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artifacts in the analysis.

@@

%Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches. Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$). %The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.
All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches. The predictive power of each one of these features is shown in the form of the Receiving Operating Characteristic (ROC) curves in Figure~\ref{fig:rocs}. The features are arranged in descending order according to the area under the curve (AUC). The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys. The maximal number of CMS is only in the seventh position here. The overall high AUC values indicate that we can successfully employ machine learning methods to distinguish matches from non-matches.
%Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3D topological surface measurements. There are quite large differences in these averages, indicating

Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files. The resulting tree is shown in Figure~\ref{fig:tree}. A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} lands is being matched correctly. Interestingly, the number of consecutive matching striae does not feature in this evaluation. 
Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree. Aside from  cross correlation, the total number of matches is also included in the decision rule. 
Between cross-correlation and CMS, cross-correlation has higher  predictive power. This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2D microscopic images of bullets.
Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure. The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}. However, in that study, authors did not consider the number of matches and non-matches.

\begin{figure}[hbtp]
  \centering
<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\textwidth'>>=
library(rpart)
library(rpart.plot)
library(RColorBrewer)
vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
names(bstats)[9:10] <- c("#matches", "#non-matches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
cc <- read.csv("csvs/crosscuts-25.csv")
# cc$bullet <- gsub(".*//","",cc$path)
# cc$bullet <- gsub(".x3p", "", cc$bullet)
# bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
#                   bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]

@@

The feasibility of creating a database of ballistic images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottom line of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks. This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system. A further re-assessment by \citet{deceuster:2015} came to the same conclusions based on the EvoFinder system. 
The NRC report also found that results from 2D images can be improved when matches are based on 3D images. This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the total number of matches and mismatches are available for a match based on 2D features.

### 2016-04-05 imaging-paper.tex
Include the arXiv id
Fix all the build issues

\end{knitrout}
\caption{\label{fig:crosscuts2}Signatures  for barrel 3, bullet~1-1 extracted from varying heights. Initially, the match between signatures taken at heights 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet. At a  level of $175\mu m$ the bullet's signature stabilizes. For this land, matches should not be attempted at lower heights. }
\end{figure}
%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
by comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures. In the approach here, we keep increasing the height $x$ at which the signature is taken until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
`stability' is  defined as two aligned signatures from heights chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.

@@

Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches. Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$). %The average distance between known matches in the Hamby study is $1.86\mu m$ with a standard deviation of $2.31\mu m$; while known mismatches have an average  distance of  $3.27\mu m$ with a standard deviation of $1.93\mu m$.
All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches. The predictive power of each one of these features is shown in the form of the Receiving Operating Characteristic (ROC) curves in Figure~\ref{fig:rocs}. The features are arranged in descending order according to the area under the curve (AUC). The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys. The maximal number of CMS is only in the seventh position here. The overall high AUC values indicate, that we can successfully employ machine learning methods to distinguish matches from non-matches.
%Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3D topological surface measurements. There are quite large differences in these averages, indicating

Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files. The resulting tree is shown in Figure~\ref{fig:tree}. A total of 132 lands is being matched correctly. Interestingly, the number of consecutive matching striae does not feature in this evaluation. 
Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree. Aside from  cross correlation, the total number of matches is also included in the decision rule. 
Between cross-correlation and CMS, cross-correlation has higher  predictive power. This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2D microscopic images of bullets.
Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure. The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}. However, in this study authors did not consider the number of matches and non-matches.

>>>>

\end{knitrout}
\caption{\label{fig:crosscuts2}Signatures  for barrel 3, bullet~1-1 extracted from varying heights. Initially, the match between signatures taken at heights 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet. At a  level of $175\mu m$ the bullet's signature stabilizes. For this land, matches should not be attempted at lower heights. }
\end{figure}
%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:

@@

%Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches. Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$). %The average distance between known matches in the Hamby study is $mean(bstats$D[bstats$match])\mu m$ with a standard deviation of $sd(bstats$D[bstats$match])\mu m$; while known mismatches have an average  distance of  $mean(bstats$D[!bstats$match])\mu m$ with a standard deviation of $sd(bstats$D[!bstats$match])\mu m$.
All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches. The predictive power of each one of these features is shown in the form of the Receiving Operating Characteristic (ROC) curves in Figure~\ref{fig:rocs}. The features are arranged in descending order according to the area under the curve (AUC). The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys. The maximal number of CMS is only in the seventh position here. The overall high AUC values indicate that we can successfully employ machine learning methods to distinguish matches from non-matches.
%Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3D topological surface measurements. There are quite large differences in these averages, indicating

Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files. The resulting tree is shown in Figure~\ref{fig:tree}. A total of 132 lands is being matched correctly. Interestingly, the number of consecutive matching striae does not feature in this evaluation. 
Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree. Aside from  cross correlation, the total number of matches is also included in the decision rule. 
Between cross-correlation and CMS, cross-correlation has higher  predictive power. This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2D microscopic images of bullets.
Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure. The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}. However, in that study, authors did not consider the number of matches and non-matches.

### 2016-04-05 imaging-supplement.Rnw
Fix all the build issues

new_processBullets <- function(bullet, x = 100) {
  crosscuts <- unique(fortify_x3p(bullet)$x)
  crosscuts <- crosscuts[crosscuts >= min(x)]
  crosscuts <- crosscuts[crosscuts <= max(x)]
  if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]

  list_of_fits <- lapply(crosscuts, function(x) {
    br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
    br111.groove <- get_grooves(br111)
    fit_loess(br111, br111.groove)$resid$data
  })
  lof <- list_of_fits %>% bind_rows

  cbind(lof, bullet = bullet$path)
}

new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
  lof <- rbind(lof1, lof2)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

  b12 <- unique(lof$bullet)
  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)

  #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
  #    theme_bw() +
  #    geom_rect(data=peaks1$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25)) +
  #    geom_rect(data=peaks2$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25))

  peaks1$lines$bullet <- b12[1]
  peaks2$lines$bullet <- b12[2]

  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)

  #   p <- qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
  #     theme_bw() +
  #     geom_rect(data=lines, aes(xmin=xmin, xmax=xmax, fill = factor(type)),  ymin=-6, ymax=6, inherit.aes = FALSE, alpha=I(0.25)) +
  #     ylim(c(-6,6)) +
  #     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match), inherit.aes = FALSE) +
  #     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match), inherit.aes = FALSE)

  maxCMS <- maxCMS(lines$match==TRUE)
  list(maxCMS = maxCMS, ccf = bAlign$ccf, lag=bAlign$lag, lines=lines, bullets=lofX)
}

@@

if (!file.exists("csvs/crosscut-sd.csv")) {
  bullets_processed <- lapply(all_bullets, function(bul) {
    cat("Computing processed bullet", basename(bul$path), "\n")
    crosscuts <- 25*(1:20)
    dframe <- new_processBullets(bullet = bul, x = crosscuts)
    dframe$bullet <- with(dframe, paste(bullet, x))
    dframe
  })
  names(bullets_processed) <- as.character(ccs$path)

@@

  write.csv(stats, "csvs/crosscut-sd.csv", row.names=FALSE)
} else {
  stats <- read.csv("csvs/crosscut-sd.csv")
}  
@
%
Figure~\ref{fig:overview} shows an overview of the signatures at different heights on a single bullet. 
\begin{figure}[hbtp]
<<one-bulletland-sd, echo=FALSE, dependson='setup-signatures', fig.height=5, fig.width=8, out.width='0.7\\textwidth', warning=FALSE>>=
k <- 1
x <- as.character(ccs$path)[k]
land <- read.x3p(x)
land[[3]] <- x
names(land)[3] <- "path"

crosscuts <- 25*(1:20)
dframe <- new_processBullets(bullet = land, x = crosscuts)
dframe$bullet <- with(dframe, paste(bullet, x))

@@

However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0. The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate. The false positive rate is therefore virtually unchanged, while we see a slight improvement in the false negative rate for an overall OOB error rate of 0.3\%, i.e.\ an increase to twice the number of comparisons leads to a decrease of 25\% of the estimated error rate. This is yet another argument in favor of a larger database for training algorithms.

@@

<<aligned, echo=FALSE, warning=FALSE,  cache=TRUE>>=
matches <- read.csv("csvs/matches.csv", header = FALSE)
matches$barrel <- rep(1:10, each=6)
matches$id <- 1:nrow(matches)
library(reshape2)
mm <- melt(matches, id.var=c("barrel", "id"), na.rm=TRUE)
mm <- subset(mm, value != "")
mm$value <- gsub("Br.* Bullet ", "", mm$value)
mm$prefix <- sprintf("Br%d Bullet ", mm$barrel)
unkns <- grep("[A-Z]",mm$value)
mm$prefix[unkns] <- "Ukn Bullet "
mm$path <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P1of2//"
mm$path[unkns] <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P2of2//"
mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))

ccs <- read.csv("csvs/crosscuts-25-new.csv")
mm <- merge(mm, ccs, by="path", all=TRUE)
mm <- subset(mm, !is.na(mm$cc))

@@

crossSmooth <- bulletSmooth(crosscuts)

@@

\begin{figure}[hbtp]
<<aligned-second, dependson='aligned', echo=FALSE, warning=FALSE, fig.width=15, fig.height=10, out.width='7.5in'>>=
qplot(x=y, y=l30, data=subset(aligned, barrel > 5), 
      geom="line", group=bullet, colour=bullet) + 
  ylim(c(-7,7)) + xlab("") + ylab("") + xlim(c(0,3000)) +
  facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
  theme(plot.margin=unit(c(0,0,-1,-1), unit="line"), legend.position="none") + 
  geom_label(aes(y=2.5*(id-3), label=label, colour=bullet), 
             x = 3000, data=subset(labels, barrel > 5), 
             inherit.aes = FALSE, hjust="right", size=3) +
  scale_colour_manual(values=cols)
@
\caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet lands for  barrels 5 to 10 of the Hamby study. }
\end{figure}
\end{landscape}
\newpage
\section{Table of feature importance}\label{supp:randomforest}
Two random forests were calculated for the Hamby study. For the first random forest only comparisons of bullet lands from known bullets and unknown bullets were used. The second random forest is based on a full comparison of every land with every other land, increasing the number of comparisons from originally 10,384 (10,212 known non-matches and 172 known matches) by another 10,931 comparisons (10,637 known non-matches and 94 known matches). Random forests allow an assessment of variable importance (also called feature importance) as the mean decrease in Gini index when including each variable. 
Table~\ref{tab:importance} shows the results for feature importance for both of these random forests. Importance~1 refers to the smaller subset, Importance~2 is the feature importance derived from the random forest based on all pairwise comparisons. 
<<setup-rtrees, echo=FALSE, message=FALSE>>=
bstats <- read.csv("data-25-25/bullet-stats.csv", stringsAsFactors = FALSE)

>>>>

@@

if (!file.exists("csvs/crosscuts-sd.csv")) {
  bullets_processed <- lapply(all_bullets, function(bul) {
    #cat("Computing processed bullet", basename(bul$path), "\n")
    crosscuts <- 25*(1:20)
    dframe <- processBullets(bullet = bul, x = crosscuts)
    dframe$bullet <- with(dframe, paste(bullet, x))
    dframe
  })
  names(bullets_processed) <- as.character(ccs$path)

@@

  write.csv(stats, "csvs/crosscuts-sd.csv", row.names=FALSE)
} else {
  stats <- read.csv("csvs/crosscuts-sd.csv")
}
@
%
Figure~\ref{fig:overview} shows an overview of the signatures at different heights on a single bullet. 
\begin{figure}[hbtp]
<<one-bulletland-sd, echo=FALSE, dependson='setup-signatures', fig.height=5, fig.width=8, out.width='0.7\\textwidth', warning=FALSE>>=
k <- 1
x <- as.character(ccs$path)[k]
land <- read.x3p(x)
land[[3]] <- x
names(land)[3] <- "path"

crosscuts <- 25*(1:20)
dframe <- processBullets(bullet = land, x = crosscuts)
dframe$bullet <- with(dframe, paste(bullet, x))

@@

However, if we use all of the available comparisons to fit another random forest of 300 trees, the defacto error rates for false positives and false negatives are again at 0. The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate. The false positive rate is therefore virtually unchanged, while we see a slight improvement in the false negative rate for an overall OOB error rate of 0.3\%, i.e.\ an increase to twice the number of comparisons leads to a decrease of 25\% of the estimated error rate. This is yet another argument in favor of a larger database for training algorithms.

@@

<<aligned, echo=FALSE, warning=FALSE,  cache=TRUE>>=
matches <- read.csv("csvs/matches.csv", header = FALSE)
matches$barrel <- rep(1:10, each=6)
matches$id <- 1:nrow(matches)
library(reshape2)
mm <- melt(matches, id.var=c("barrel", "id"), na.rm=TRUE)
mm <- subset(mm, value != "")
mm$value <- gsub("Br.* Bullet ", "", mm$value)
mm$prefix <- sprintf("Br%d Bullet ", mm$barrel)
unkns <- grep("[A-Z]",mm$value)
mm$prefix[unkns] <- "Ukn Bullet "
mm$path <- "app/images/Hamby252_3DX3P1of2//"
mm$path[unkns] <- "app/images/Hamby252_3DX3P2of2//"
mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))

ccs <- read.csv("csvs/crosscuts-25.csv")
mm <- merge(mm, ccs, by="path", all=TRUE)
mm <- subset(mm, !is.na(mm$cc))

@@

crosscuts$bullet <- crosscuts$path
crossSmooth <- bulletSmooth(crosscuts)

@@

\begin{figure}[hbtp]
<<aligned-second, dependson='aligned', echo=FALSE, warning=FALSE, fig.width=15, fig.height=10, out.width='7.5in'>>=
qplot(x=y, y=l30, data=subset(aligned, barrel > 5), 
      geom="line", group=bullet, colour=bullet) + 
  ylim(c(-7,7)) + xlab("") + ylab("") + xlim(c(0,3000)) +
  facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
  theme(plot.margin=unit(c(0,0,-1,-1), unit="line"), legend.position="none") + 
  geom_label(aes(y=2.5*(id-3), label=label, colour=bullet), 
             x = 3000, data=subset(labels, barrel > 5), 
             inherit.aes = FALSE, hjust="right", size=3) +
  scale_colour_manual(values=cols)
@
\caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet lands for  barrels 5 to 10 of the Hamby study. }
\end{figure}
\end{landscape}
\newpage
\section{Table of feature importance}\label{supp:randomforest}
Two random forests were calculated for the Hamby study. For the first random forest only comparisons of bullet lands from known bullets and unknown bullets were used. The second random forest is based on a full comparison of every land with every other land, increasing the number of comparisons from originally 10,384 (10,212 known non-matches and 172 known matches) by another 10,931 comparisons (10,637 known non-matches and 94 known matches). Random forests allow an assessment of variable importance (also called feature importance) as the mean decrease in Gini index when including each variable. 
Table~\ref{tab:importance} shows the results for feature importance for both of these random forests. Importance~1 refers to the smaller subset, Importance~2 is the feature importance derived from the random forest based on all pairwise comparisons. 
<<setup-rtrees, echo=FALSE, message=FALSE>>=
bstats <- read.csv("data-25-25/bullet-stats-old.csv", stringsAsFactors = FALSE)
